---
title: "P8105_hw2_my2838"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

# Problem 1

First, clean the data in pols-month.csv.

```{r}
pols_df = 
  read.csv("data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(col = mon, into = c("year","month","day"), sep = "-") |>
  mutate(
    year = as.integer(year),
    month = month(as.integer(month),label = TRUE),
    gov_president = gov_gop + gov_dem,
    sen_president = sen_gop + sen_dem,
    rep_president = rep_gop + rep_dem) |>
  mutate(
    date = paste(year, month, sep="-")
  ) |>
  select(-ends_with("dem"),-ends_with("gop"),-day,-year,-month)
```

Second, clean the data in snp.csv.

```{r}
snp_df = 
  read.csv("data/fivethirtyeight_datasets/snp.csv") |>
  separate(col = date, into = c("month","day","year"), sep = "/") |>
  mutate(year = as.integer(year)) |>
  mutate(year = ifelse(year > 15, year + 1900, year + 2000),
         month = month(as.integer(month),label = TRUE),
         date = paste(year, month, sep="-")) |>
  arrange(year, month) |>
  select(-day, -year, -month)
```

Third, tidy the unemployment data.

```{r}
unemployment_df = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv",show_col_types = FALSE) |>
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "unemployment"
  ) |>
  mutate(
    year = as.integer(Year),
    month = as.character(month),
    unemployment = as.double(unemployment),
    date = paste(year, month, sep="-")
  ) |>
  select(-Year,-year,-month)
```

```{r}
final_df = 
  full_join(pols_df, snp_df, by = "date") |>
  full_join(unemployment_df, by = "date") |>
  separate(date, into = c("year","month"), sep = "-") |>
  select( year, month, everything())
```


# Problem 2

Read and clean the Mr. Trash Wheel sheet.

```{r}
mr_trash_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", skip = 1, range = cell_cols('A:N')) |>
  janitor::clean_names()  |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    wheel = "mr")
```

Use a similar process to import, clean, and organize the data for Professor Trash Wheel and Gwynnda.

```{r}
pro_trash_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", skip = 1, range = cell_cols('A:M')) |>
  janitor::clean_names()  |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    wheel = "pro")

gwy_trash_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", skip = 1, range = cell_cols('A:K')) |>
  janitor::clean_names()  |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    wheel = "gwy")
```

Combine these dataframe, and add an additional variable `homes_powered`.

```{r}
trash_df = 
  bind_rows(mr_trash_df, pro_trash_df, gwy_trash_df) |>
  mutate(homes_powered = 500* weight_tons /30 ) 
```

*The combined master data sets for trash wheel has `r nrow(trash_df)` rows and `r ncol(trash_df)` columns. *

*The total weight of trash collected by Professor trash wheel is `r pull(pro_trash_df,weight_tons) |> sum()` tons.*

*The total number of cigarette butts collected by Gwynnda in July of 2021 is `r filter(gwy_trash_df, year == 2021 & month == "July") |> pull(cigarette_butts) |> sum() `.*



# Problem 3

Import, clean, and tidy the dataset of baseline demographics.

```{r}

baseline =
  read_csv("data/data_mci/MCI_baseline.csv", skip = 1, show_col_types = FALSE) |>
  janitor::clean_names() |>
  filter(current_age < age_at_onset | age_at_onset == ".") |>
  mutate(
      sex = case_match(
      sex,
      1 ~ "male",
      0 ~ "female"),
      apoe4 = as.logical(apoe4),
      current_age = as.numeric(current_age),
      age_at_onset = as.numeric(age_at_onset)
  ) 
```

*There are `r nrow(baseline)` participants were recruited, and `r  drop_na(baseline) |> nrow() ` develop MCI.*

*The average baseline age is `r pull(current_age) |> mean()`*

*The proportion of women in the study are APOE4 carriers is `r nrow(subset(baseline, sex == "female" & apoe4 == 1))/nrow(subset(baseline, sex == "female"))`*



Import, clean, and tidy the dataset of longitudinally observed biomarker values.

```{r}
amyloid =
  read_csv("data/data_mci/mci_amyloid.csv", skip = 1, show_col_types = FALSE) |>
  janitor::clean_names() |>
  drop_na(baseline) |>
  rename("id" = "study_id")
```

*When import this dataset, I need to drop the NA in baseline, and rename the variable "Study Id" to further match the same variable in baseline dataset.*



Combine two datasets to keep all the participants appear.

```{r}
full_data = full_join(amyloid, baseline, by = "id")
```

Combine two datasets to keep participants who only appear in both datasets.

```{r}
inn_data = inner_join(amyloid, baseline, by = "id")
```


*We have `r nrow(full_data)` participants in the research. But only `r nrow(inn_data)` participants appear in both datasets, which means there are `r nrow(full_data)-nrow(inn_data)` participants appear in only the baseline or amyloid datasets.*

*The resulting dataset of participants who appear in both baseline and amyloid datasets have `r nrow(inn_data)` rows and `r ncol(inn_data)` columns.*



Let's save the result.

```{r}
write.csv(inn_data, file="both_amyloid_baseline.csv")
```

